---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  warning = FALSE,
  message = FALSE,
  cache = TRUE
)
```

[![Travis-CI Build Status](https://travis-ci.org/jjchern/qhp.svg?branch=master)](https://travis-ci.org/jjchern/qhp)


# About

This package includes Qualified Health Plan (qhp) data in the Health Insurance Marketplace. So far it contains three years of enrollment data and the initial health plan data.

# Installation

```r
# install.packages("devtools")
devtools::install_github("jjchern/qhp")
```

# Links

## General

- [ASPE AFFORDABLE CARE ACT RESEARCH](http://aspe.hhs.gov/affordable-care-act-research)

## Sources

- [PLAN SELECTIONS BY ZIP CODE IN THE HEALTH INSURANCE MARKETPLACE, 2014](http://aspe.hhs.gov/plan-selections-zip-code-health-insurance-marketplace-september-2014)
- [PLAN SELECTIONS BY ZIP CODE IN THE HEALTH INSURANCE MARKETPLACE, 2015](http://aspe.hhs.gov/plan-selections-zip-code-health-insurance-marketplace-april-2015)
- [PLAN SELECTIONS BY ZIP CODE AND COUNTY IN THE HEALTH INSURANCE MARKETPLACE: MARCH 2016](https://aspe.hhs.gov/basic-report/plan-selections-zip-code-and-county-health-insurance-marketplace-march-2016)
- [HEALTH PLAN DATASETS](https://www.healthcare.gov/health-and-dental-plan-datasets-for-researchers-and-issuers/)

## Inital Enrollment Data

- [HEALTH INSURANCE MARKETPLACE: SUMMARY ENROLLMENT REPORT FOR THE INITIAL ANNUAL OPEN ENROLLMENT PERIOD](https://aspe.hhs.gov/report/health-insurance-marketplace-summary-enrollment-report-initial-annual-open-enrollment-period)
- [ADDENDUM TO THE HEALTH INSURANCE MARKETPLACE SUMMARY ENROLLMENT REPORT FOR THE INITIAL ANNUAL OPEN ENROLLMENT PERIOD](https://aspe.hhs.gov/sites/default/files/pdf/93351/ib_2014Apr_enrollAddendum.pdf)

## Second Enrollment Data

- [QHP Selections by Age Groups](https://data.healthcare.gov/dataset/2015-Qualifying-Health-Plan-Selections-by-Age-Grou/qv54-6mtm)
- [QHP Selections by APTC](https://data.healthcare.gov/dataset/2015-Qualifying-Health-Plan-Selections-by-APTC-and/f7yt-nc7m)
- [QHP Selections by Household Incomes](https://data.healthcare.gov/dataset/2015-Qualifying-Health-Plan-Selections-by-Househol/sy6b-59wj)
- [QHP Selections by Race/ethnicity](https://data.healthcare.gov/dataset/2015-Qualifying-Health-Plan-Selections-by-Race-Eth/dvpf-jb7v)
- [QHP Selections by Types of Selections](https://data.healthcare.gov/dataset/2015-Qualifying-Health-Plan-Selections-by-Type-of-/kjvw-ka5c)
- [QHP Selections by CSR](https://data.healthcare.gov/dataset/2015-Qualifying-Health-Plan-Selections-by-CSR-and-/ncbw-vdsn)
- [QHP Selections by Metal Levels](https://data.healthcare.gov/dataset/2015-Qualifying-Health-Plan-Selections-by-Metal-Le/d89b-9897)

# Usage

## Zip Code Level Enrollment Data

```{r, message=FALSE, warning=FALSE}
library(dplyr, warn.conflicts = FALSE)
qhp::enrollment2014
qhp::enrollment2015
qhp::enrollment2016
```

## Top 10 Counties, Sorted According to Total Enrollment

```{r, message=FALSE, warning=FALSE}
# 2014
qhp::enrollment2014 %>% 
  na.omit() %>% 
  group_by(countygeoid) %>% 
  summarise(countyname = first(countyname),
            enrollment = sum(PlanSelections),
            state = first(state),
            copop = first(copop)) %>% 
  arrange(desc(enrollment))

# 2015
qhp::enrollment2015 %>% 
  na.omit() %>% 
  group_by(countygeoid) %>% 
  summarise(countyname = first(countyname),
            enrollment = sum(PlanSelections),
            state = first(state),
            copop = first(copop)) %>% 
  arrange(desc(enrollment))
```

## Top 10 Counties, Sorted According to Enrollment per 10,000 Residents

```{r, message=FALSE, warning=FALSE}
# 2014
qhp::enrollment2014 %>% 
  na.omit() %>% 
  group_by(countygeoid) %>% 
  summarise(countyname = first(countyname),
            enrollment = sum(PlanSelections),
            state = first(state),
            copop = first(copop)) %>% 
  mutate(enroll_per100000 = enrollment / copop * 10000) %>% 
  filter(copop >= 100000) %>% 
  arrange(desc(enroll_per100000))

# 2015
qhp::enrollment2015 %>% 
  na.omit() %>% 
  group_by(countygeoid) %>% 
  summarise(countyname = first(countyname),
            enrollment = sum(PlanSelections),
            state = first(state),
            copop = first(copop)) %>% 
  mutate(enroll_per100000 = enrollment / copop * 10000) %>% 
  filter(copop >= 100000) %>% 
  arrange(desc(enroll_per100000))
```

## Enrolloment per 10,000 Residents by counties

```{r county-map, echo=FALSE}
# 2014
qhp::enrollment2014 %>% 
  na.omit() %>% 
  group_by(countygeoid) %>% 
  summarise(countyname = first(countyname),
            enrollment = sum(PlanSelections),
            state = first(state),
            copop = first(copop)) %>% 
  mutate(enroll_per10000 = enrollment / copop * 10000) %>% 
  arrange(desc(enroll_per10000)) %>% 
  mutate(countygeoid = sprintf("%05.0f", countygeoid)) %>% 
  arrange(countygeoid) %>% 
  mutate(year = 2014) -> en_county_2014

# 2015
qhp::enrollment2015 %>% 
  na.omit() %>% 
  group_by(countygeoid) %>% 
  summarise(countyname = first(countyname),
            enrollment = sum(PlanSelections),
            state = first(state),
            copop = first(copop)) %>% 
  mutate(enroll_per10000 = enrollment / copop * 10000) %>% 
  arrange(desc(enroll_per10000)) %>% 
  mutate(countygeoid = sprintf("%05.0f", countygeoid)) %>% 
  arrange(countygeoid) %>% 
  mutate(year = 2015) -> en_county_2015

# 2016
qhp::enrollment2016 %>% 
  na.omit() %>% 
  group_by(countygeoid) %>% 
  summarise(countyname = first(countyname),
            enrollment = sum(PlanSelections),
            state = first(state),
            copop = first(copop)) %>% 
  mutate(enroll_per10000 = enrollment / copop * 10000) %>% 
  arrange(desc(enroll_per10000)) %>% 
  mutate(countygeoid = sprintf("%05.0f", countygeoid)) %>% 
  arrange(countygeoid) %>% 
  mutate(year = 2016) -> en_county_2016

# devtools::install_github("jjchern/usmapdata")
usmapdata::county %>% 
  left_join(en_county_2014, by = c("id" = "countygeoid")) %>% 
  mutate(region = id) -> en_county_2014_map

usmapdata::county %>% 
  left_join(en_county_2015, by = c("id" = "countygeoid")) %>% 
  mutate(region = id) -> en_county_2015_map

usmapdata::county %>% 
  left_join(en_county_2016, by = c("id" = "countygeoid")) %>% 
  mutate(region = id) -> en_county_2016_map

en_county_map = bind_rows(
  en_county_2014_map, 
  en_county_2015_map,
  en_county_2016_map
) %>% 
  filter(enroll_per10000 < 3000)

usmapdata::state %>% 
  mutate(region = id) -> state_map

library(ggplot2)

ggplot() +
  geom_map(data = en_county_2014_map, map = en_county_2014_map,
           aes(x = long, y = lat, map_id = region, fill = enroll_per10000),
           colour = alpha("white", 0.1), size=0.2) +
  geom_map(data = state_map, map = state_map,
           aes(x = long, y = lat, map_id = region),
           colour = "white", fill = "NA") +
  coord_map("albers", lat0 = 30, lat1 = 40) +
  scale_fill_gradientn(colors = viridis::viridis(en_county_map$enroll_per10000, option = "B"),
                       limits=c(min(en_county_map$enroll_per10000, na.rm = TRUE),
                                max(en_county_map$enroll_per10000, na.rm = TRUE))) +
  ggtitle("Enrollment per 10,000 Population in 2014") +
  ggthemes::theme_map() +
  theme(legend.position = c(.85, .3),
        legend.title=element_blank())

ggplot() +
  geom_map(data = en_county_2015_map, map = en_county_2015_map,
           aes(x = long, y = lat, map_id = region, fill = enroll_per10000),
           colour = alpha("white", 0.1), size=0.2) +
  geom_map(data = state_map, map = state_map,
           aes(x = long, y = lat, map_id = region),
           colour = "white", fill = "NA") +
  coord_map("albers", lat0 = 30, lat1 = 40) +
  scale_fill_gradientn(colors = viridis::viridis(en_county_map$enroll_per10000, option = "B"),
                       limits=c(min(en_county_map$enroll_per10000, na.rm = TRUE),
                                max(en_county_map$enroll_per10000, na.rm = TRUE))) +
  ggtitle("Enrollment per 10,000 Population in 2015") +
  ggthemes::theme_map() +
  theme(legend.position = c(.85, .3),
        legend.title=element_blank())

ggplot() +
  geom_map(data = en_county_2016_map, map = en_county_2016_map,
           aes(x = long, y = lat, map_id = region, fill = enroll_per10000),
           colour = alpha("white", 0.1), size=0.2) +
  geom_map(data = state_map, map = state_map,
           aes(x = long, y = lat, map_id = region),
           colour = "white", fill = "NA") +
  coord_map("albers", lat0 = 30, lat1 = 40) +
  scale_fill_gradientn(colors = viridis::viridis(en_county_map$enroll_per10000, option = "B"),
                       limits=c(min(en_county_map$enroll_per10000, na.rm = TRUE),
                                max(en_county_map$enroll_per10000, na.rm = TRUE))) +
  ggtitle("Enrollment per 10,000 Population in 2016") +
  ggthemes::theme_map() +
  theme(legend.position = c(.85, .3),
        legend.title=element_blank())
```


## Total enrollment by states

```{r, message=FALSE, warning=FALSE}
# 2014
qhp::enrollment2014 %>% 
  na.omit() %>% 
  group_by(state) %>% 
  summarise(statename = first(StateName),
            enrollment = sum(PlanSelections)) %>% 
  arrange(enrollment %>% desc) %>% 
  print(n = 36)

# 2015
qhp::enrollment2015 %>% 
  na.omit() %>% 
  group_by(state) %>% 
  summarise(statename = first(StateName),
            enrollment = sum(PlanSelections)) %>% 
  arrange(enrollment %>% desc) %>% 
  print(n = 37)
```

## Health Plan Data

```{r, message=FALSE, warning=FALSE}
qhp::qhp2014
```

# Reference

- [Summary of the Affordable Care Act](http://kff.org/health-reform/fact-sheet/summary-of-the-affordable-care-act/)

- [42 U.S. Code Subchapter IV - AFFORDABLE COVERAGE CHOICES FOR ALL AMERICANS](https://www.law.cornell.edu/uscode/text/42/chapter-157/subchapter-IV)
    - [PART A - Premium Tax Credits and Cost-Sharing Reductions (§ 18071)](https://www.law.cornell.edu/uscode/text/42/chapter-157/subchapter-IV/part-A)
    - [PART B - Eligibility Determinations (§§ 18081 to 18084)](https://www.law.cornell.edu/uscode/text/42/chapter-157/subchapter-IV/part-B)

- [42 U.S. Code Part A - Establishment of Qualified Health Plans](https://www.law.cornell.edu/uscode/text/42/chapter-157/subchapter-III/part-A)
    - [42 U.S. Code § 18021 - Qualified health plan defined](https://www.law.cornell.edu/uscode/text/42/18021)
    - [42 U.S. Code § 18022 - Essential health benefits requirements](https://www.law.cornell.edu/uscode/text/42/18022)

- [Health Insurance Market Reforms](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/index.html)
    - [Annual Limits](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/Annual-Limits.html)
    - [Coverage for Young Adults](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/Coverage-for-Young-Adults.html)
    - [Grandfathered Plans](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/Grandfathered-Plans.html)
    - [Market Rating Reforms](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/Market-Rating-Reforms.html)
        - [**Geographic Rating Areas**](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/state-gra.html)
        - [**Other State Specific Rating Variations**](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/state-rating.html)
    - [Medical Loss Ratio](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/Medical-Loss-Ratio.html)
    - [Minimum Essential Coverage](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/minimum-essential-coverage.html)
    - [Patient's Bill of Rights](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/Patients-Bill-of-Rights.html)
    - [Prevention](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/Prevention.html)
    - [Review of Insurance Rates](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/Review-of-Insurance-Rates.html)
        - [**Rate Review Grants**](https://www.cms.gov/CCIIO/Resources/Rate-Review-Grants/index.html)
            - [Overview of the health insurance premium grant program](https://www.cms.gov/CCIIO/Resources/Regulations-and-Guidance/rates.html)
    - [Student Health Plans](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/Student-Health-Plans.html)
    - [Self-Funded, Non-Federal Governmental Plans](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/nonfedgovplans.html)
    - [Compliance](https://www.cms.gov/CCIIO/Programs-and-Initiatives/Health-Insurance-Market-Reforms/compliance.html)
